Figure 2 and figure 2 supplementary material

### 2022-02-11 Figures/Figure2.Rmd
wrap tumour purity scatterplot + fix ridgeplot text adjust + loop through cancers

    ct_subset = data %>%
      dplyr::select(matches(ct)) %>%
      rename("H&E" = ct) %>%
      gather() %>%
      mutate(key = str_remove_all(key, paste(substring_remove, collapse = "|")),
             key=factor(key, levels=names(cor.ct.sorted)),
             method = as.factor(str_extract(key, "[^_]+")))

  final_plot[[ct]] <- ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = 1, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type")
  }
  return(final_plot)
}

@@

@@

# Figure
## A) Tumor Purity
```{r}
Aran_all_facs %>% dplyr::filter(Cancer_type == "LUAD", CPE != "NaN") %>% 
  arrange(factor(Sample_ID, levels = Sample_ID))  -> Facs_LUAD_

Comm_samples <- Reduce(intersect, list(tcga_deconv$Samples,
                                       Facs_LUAD_$Sample_ID))
Facs_LUAD_ %>% 
  dplyr::filter(Sample_ID %in% Comm_samples) %>% 
  arrange(factor(Sample_ID, levels = Comm_samples)) -> Facs_LUAD

tcga_deconv %>% 
  dplyr::filter(Samples %in% Comm_samples) %>% 
  arrange(factor(Samples, levels = Comm_samples)) -> LUAD_RNA_decon

# Scatterplot CCLE_TIL10 LUAD
identical(Facs_LUAD$Sample_ID, LUAD_RNA_decon$Samples) %>% stopifnot()
 Facs_LUAD %>% bind_cols(LUAD_RNA_decon) -> LUAD_merged

title = "ESTIMATE vs CCLE TIL10 Cancer"
color = "black"
scatterplot_truevsestimate(LUAD_merged, "ESTIMATE", "Epidish_CCLE_TIL10_Cancer.cells", title, color)

# Scatterplot BPRNACan LUAD
title = "ESTIMATE vs BPRNACan Cancer"
color = "black"
scatterplot_truevsestimate(LUAD_merged, "ESTIMATE", "Epidish_BPRNACan_cancer", title, color)
```

@@

ridgeplots_by_cancer <- list()
for (cancer in unique(HE_all_TCGA$Study)){
ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
}

# pdf_name <- paste("test", "ridges", ".pdf", sep = "")
#   ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 2)
#

>>>>

    ct_subset = data %>%
      dplyr::select(matches(ct)) %>%
      rename("H&E" = ct) %>%
      gather() %>%
      mutate(key = str_remove_all(key, paste(substring_remove, collapse = "|")),
             key=factor(key, levels=names(cor.ct.sorted)),
             method = as.factor(str_extract(key, "[^_]+")))
  adjust_x <- max(ct_subset$value)
  final_plot[[ct]] <- print(ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = adjust_x, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_x_continuous(limits = c(0, (adjust_x + adjust_x*0.2))) +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type"))
  }
  return(final_plot)
}

@@

wrapper_tpurity <- function(deconv_data, true_estimates, cancer_abreviation){
  true_estimates %>% dplyr::filter(Cancer_type == cancer_abreviation, CPE != "NaN") %>% 
    arrange(factor(Sample_ID, levels = Sample_ID))  -> true_subset_

  Comm_samples <- Reduce(intersect, list(deconv_data$Samples,
                                         true_subset_$Sample_ID))
  true_subset_ %>% 
    dplyr::filter(Sample_ID %in% Comm_samples) %>% 
    arrange(factor(Sample_ID, levels = Comm_samples)) -> true_subset

  # Scatterplot CCLE_TIL10
  identical(true_subset$Sample_ID, deconv_subset$Samples) %>% stopifnot()
   true_subset %>% bind_cols(deconv_subset) -> merged_subset

  final_plot = list()
  color = "black"

  title = "ESTIMATE vs CCLE TIL10 Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_CCLE_TIL10_Cancer.cells", title, color)

  # Scatterplot BPRNACan
  title = "ESTIMATE vs BPRNACan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, "ESTIMATE", "Epidish_BPRNACan_cancer", title, color)

  return(final_plot)
}

@@

# Figure
## A) Tumor Purity
```{r}
wrapper_tpurity(deconv_data = tcga_deconv,true_estimates = Aran_all_facs, cancer_abreviation = "LUAD")
```

@@

ridgeplots_by_cancer <- list()
tpurity_by_cancer <- list()
blank  <- ggplot() + theme_void()

for (cancer in unique(HE_all_TCGA$Study)){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank
}

ridges <- ggarrange(plotlist = ridgeplots_by_cancer[["LUAD"]], ncol = 3, common.legend = T, legend = "bottom")
scater <- ggarrange(plotlist = tpurity_by_cancer[["LUAD"]], ncol = 3, common.legend = T)
blank  <- ggplot() + theme_void()
fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
annotate_figure(fullplot, top = text_grob("LUAD", color = "black", size = 30))
pdf_name <- paste("test", "ridges", ".pdf", sep = "")
ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 3)
# for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
#    annotate_figure(print(ridgeplots_by_cancer[[cancer]][[ct]]),
#                              top = text_grob("Test", color = "black", size = 30),
#                              left = text_grob("Estimated fractions", color = "black", size = 20, rot = 90),
#                              bottom = text_grob("H&E estimates", color = "black", size = 20))
# }

#
